---
title: "Bayesian Logistic Regression"
author: "Alex B.Dunbar"
date: "8/9/2021"
output:
  html_document: default
  pdf_document: default
---

Clear environment, graphics and console
```{r}
# Clear environment ####
rm(list = ls()) 

# Clear plots
graphics.off()  # Clears plots, closes all graphics devices

# Clear console
cat("\014")  # ctrl+L
```
Install required packages
```{r include=FALSE}
# Install pacman ("package manager") if needed
if (!require("pacman")) install.packages("pacman")

# Load contributed packages with pacman
pacman::p_load(pacman, rio, tidyverse, GGally, magrittr, broom, skimr)  #party, 
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 1. Introduction
This is a Bayesian logistic regression related to red variants of the Portuguese "Vinho Verde" wine.  There are 11 input variables (covariates) based on physiochemical tests.  The output variable is `quality` and is a score between 0 and 10 which is based on sensory data.


# 2. Import data.
The dataset is described in the publication by Cortez, P., Cerdeira, A., Almeida, F., Matos, T., & Reis, J. (2009). [Modeling wine preferences by data mining from physicochemical properties.](https://www.sciencedirect.com/science/article/pii/S0167923609001377?casa_token=3l0RostJJXAAAAAA:QC5SBNWeoP2No8AVCA8EMu4R9yLCBh5gyZvsEMEVT1DxrzPwybfUyU0fWb8u4sqeqSyIUQBuIQI).
```{r}
wine <- read.csv("winequality-red.csv")
head(wine)
```
## Check for missing values
There are no `NA` values in the dataset.
```{r}
any(is.na(wine))
```
# 3. Classify "good" wines
We want to implement a logistic regression, therefore we want a response variable which assumes values either 0 or 1. Suppose we consider "good" a wine with quality above 6.5 (included).

We add a column called `good` that is `1` if the `quality` is greater than 6.5 and `0` if `quality` is less than 6.5.

We also add a column of strings called `drinkable`, relating 1 and 0 to "good" and "bad" as appropriate.
```{r}
 wine %<>%
  mutate(good = ifelse(quality > 6, 1, 0)) %>%
  mutate(drinkable = ifelse(quality > 6, "good", "bad")) %>%
  mutate(quality = factor(quality, levels=1:10))
```
Remove spaces from names to make references easier.
```{r}
colnames(wine) <- c("fixed_acidity", "volatile_acidity", "citric_acid",
                      "residual_sugar", "chlorides", "free_sulfur_dioxide",
                      "total_sulfur_dioxide", "density", "pH", "sulphates",
                      "alcohol", "quality", "good", "drinkable")
```
The structure of the dataset, `str(wine)`, will confirm that there are no `NA` values, the feature names, the feature type (num, Factor, chr) and the first 10 values.
```{r}
str(wine)
```
# 4. Frequentist Logistic Regression

The logistic model is computed using the generalised linear model (glm) with `good` being a function of all other covariates except `quality` and `drinkable`.

```{r}
fit <- glm(
  good ~ . - quality - drinkable,
  data = wine,
  family = binomial(link="logit")
)
```
## Significant coefficients
The significant coefficients are labelled in the last column of the summary() function.  Most significant coefficients are:

* `sulphates` (Pr($\alpha$ > 6.924 = 4.39e-12)),
* `alcohol`(Pr($\alpha$ > 5.724 = 1.04e-08)),
* `volatile_acidity` (Pr($\alpha$ > |-3.291| = 9.99e-4)), and
* `total_sulfur_dioxide` (Pr($\alpha$ > |-3.378| = 7.31e-4))

```{r}
fit %>% summary()
```
# 5. Impact on outcome by varying total_sulfur_dioxide
By fixing all coefficients and varying one, we can see the impact of one particular coefficient has on the probabilistic outcome of the logistic regression. 

As a comparison I have fixed all significant coefficients in turn (`total_sulfur_dioxide`, `volatile_acidity`, `sulphates` and `alcohol`)

First, we save coefficients to their own variable.
```{r}
b0 <- fit$coefficients[1]  # Intercept = 242.76251933
b1 <- fit$coefficients[2]  # fixed_acidity = 0.27495289
b2 <- fit$coefficients[3]  # volatile_acidity = -2.58100211
b3 <- fit$coefficients[4]  # citric_acid = 0.56779433
b4 <- fit$coefficients[5]  # residual_sugar = 0.23946420
b5 <- fit$coefficients[6]  # chlorides = -8.81636544
b6 <- fit$coefficients[7]  # free_sulfur_dioxide = 0.01082060
b7 <- fit$coefficients[8]  # total_sulfur_dioxide = -0.01653061
b8 <- fit$coefficients[9]  # density = -257.79757874
b9 <- fit$coefficients[10]  # pH = 0.22418522
b10 <- fit$coefficients[11]  # sulphates = 3.74987886
b11 <- fit$coefficients[12]  # alcohol = 0.75333905
```
Then save each individual mean values
```{r}
fixed_acidity_mean <- mean(wine$fixed_acidity)
volatile_acidity_mean <- mean(wine$volatile_acidity)
citric_acid_mean <- mean(wine$citric_acid)
residual_sugar_mean <- mean(wine$residual_sugar)
chlorides_mean <- mean(wine$chlorides)
free_sulfur_dioxide_mean <- mean(wine$free_sulfur_dioxide)
total_sulfur_dioxide_mean <- mean(wine$total_sulfur_dioxide)
density_mean <- mean(wine$density)
pH_mean <- mean(wine$pH)
sulphates_mean <- mean(wine$sulphates)
alcohol_mean <- mean(wine$alcohol)
```
Compute the range of each significant coefficient, `total_sulfur_dioxide`, `sulphates`, `alcohol`, and `volatile_acidity`
```{r}
total_sulfur_dioxide_range <- seq(from=min(wine$total_sulfur_dioxide), to=max(wine$total_sulfur_dioxide), by=1)
sulphates_range <- seq(from=min(wine$sulphates), to=max(wine$sulphates), by=0.002)
alcohol_range <- seq(from=min(wine$alcohol), to=max(wine$alcohol), by=0.05)
volatile_acidity_range <- seq(from=min(wine$volatile_acidity), to=max(wine$volatile_acidity), by=0.002)
```
Plot of the range histogram of each of the four most significant covariates.
```{r}
par(mfrow=c(2,2))
hist(total_sulfur_dioxide_range)
hist(sulphates_range)
hist(alcohol_range)
hist(volatile_acidity_range)
```

Calculate probabilities for each significant coefficient
```{r}
total_sulfur_dioxide_GOOD <- b0 + b1*fixed_acidity_mean + b2*volatile_acidity_mean +
  b3*citric_acid_mean + b4*residual_sugar_mean + b5*chlorides_mean + b6*free_sulfur_dioxide_mean +
  b7*total_sulfur_dioxide_range + b8*density_mean + b9*pH_mean + b10*sulphates_mean + b11*alcohol_mean

sulphates_GOOD <- b0 + b1*fixed_acidity_mean + b2*volatile_acidity_mean +
  b3*citric_acid_mean + b4*residual_sugar_mean + b5*chlorides_mean + b6*free_sulfur_dioxide_mean +
  b7*total_sulfur_dioxide_mean + b8*density_mean + b9*pH_mean + b10*sulphates_range + b11*alcohol_mean

alcohol_GOOD <- b0 + b1*fixed_acidity_mean + b2*volatile_acidity_mean +
  b3*citric_acid_mean + b4*residual_sugar_mean + b5*chlorides_mean + b6*free_sulfur_dioxide_mean +
  b7*total_sulfur_dioxide_mean + b8*density_mean + b9*pH_mean + b10*sulphates_mean + b11*alcohol_range

volatile_acidity_GOOD <- b0 + b1*fixed_acidity_mean + b2*volatile_acidity_range +
  b3*citric_acid_mean + b4*residual_sugar_mean + b5*chlorides_mean + b6*free_sulfur_dioxide_mean +
  b7*total_sulfur_dioxide_mean + b8*density_mean + b9*pH_mean + b10*sulphates_mean + b11*alcohol_mean
```
Histograms of what????
```{r}
par(mfrow=c(2,2))
hist(total_sulfur_dioxide_GOOD)
hist(sulphates_GOOD)
hist(alcohol_GOOD)
hist(volatile_acidity_GOOD)
```

Calculate logistic probabilities for each
```{r}
total_sulfur_dioxide_probs <- exp(total_sulfur_dioxide_GOOD)/(1 + exp(total_sulfur_dioxide_GOOD))
sulphates_probs <- exp(sulphates_GOOD) / (1 + exp(sulphates_GOOD))
alcohol_probs <- exp(alcohol_GOOD) / (1 + exp(alcohol_GOOD))
volatile_acidity_probs <- exp(volatile_acidity_GOOD) / (1 + exp(volatile_acidity_GOOD))
```
Plot logistic probabilities
```{r}
par(mfrow=c(2,2))
hist(total_sulfur_dioxide_probs)
hist(sulphates_probs)
hist(alcohol_probs)
hist(volatile_acidity_probs)
```

## Plot the results.

The plots of each of the four covariates are the probabilities across the range of 
```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(total_sulfur_dioxide_range, total_sulfur_dioxide_probs, 
     ylim=c(0,1), type="l", lwd=3, lty=2, col="gold", 
     xlab="Total Sulfur Dioxide", ylab="P(conversion)", main="Probability of conversion")

plot(sulphates_range, sulphates_probs, 
     ylim=c(0,1), type="l", lwd=3, lty=2, col="gold", 
     xlab="Total Sulphates", ylab="P(conversion)", main="Probability of conversion")

plot(alcohol_range, alcohol_probs, 
     ylim=c(0,1), type="l", lwd=3, lty=2, col="gold", 
     xlab="Total Alcohol", ylab="P(conversion)", main="Probability of conversion")

plot(volatile_acidity_range, volatile_acidity_probs, 
     ylim=c(0,1), type="l", lwd=3, lty=2, col="gold", 
     xlab="Total Volatile Acidity", ylab="P(conversion)", main="Probability of conversion")
```

# 6. Bayesian Logistic Regression

The following is a Bayesian logistic regression analysis. 

To perform a Bayesian analysis via Metropolis-Hastings algorithm we write a function defining the target distribution of the Beta coefficients.  We work in terms of log posterior to avoid numerical problems with computer computation.

## Likelihood.
The likelihood is a product of Bernoulli likelihoods for each combination of covariate and response variable for the i-th term.
$$ L(\beta; y, x) = \prod_{i=1}^n \bigg(\frac{\text{exp}(x_i \beta)}{1 + \text{exp}(x_i \beta)}\bigg)^{y_i}\bigg(1- \frac{\text{exp}(x_i \beta)}{1 + \text{exp}(x_i \beta)}\bigg)^{1-y_i} $$
## Log-likelihood
$$ \log L (\beta; y, x) = \sum_{i=1}^n y_i(x_i \beta) - \log [1 + \text{exp}(x_i \beta)] $$

## Prior
Independent normal prior distributions for each $\beta_j, j=0,1,\dots,k$, such that $\beta_j \sim N(0, 100)$

## Log posterior distribution
```{r}
lpost.LR <- function(beta, x, y) {
  # beta: vector of coefficients
  # x: covariates (regressors)
  # y: response variable
  
  # eta is the matrix product of the covariates and the coefficients
  eta <- as.numeric(x %*% beta)
  # logp: y=1, in terms of logistic function
  logp <- eta - log(1 + exp(eta))  # in log scale
  # logq: y=0, expressed in terms of logistic function
  logq <- log(1 - exp(logp))
  # sum of both contributions: when y=1 + when y=0
  logl <- sum(logp[y==1]) + sum(logq[y==0])
  # log prior: normal indep prior dists $N(0, 100)$
  lprior <- sum(dnorm(beta, 0, 10,log = T))  # 10: mixed, 50 var, 100 large var, 20 like 10, 5 low
  # return log posterior (likelihood + prior)
  return(logl + lprior)
}
```
The MLE for each coefficient can be used to initialise the simulation matrix
```{r}
fit <- glm(
  #good ~ volatile_acidity + total_sulfur_dioxide + sulphates + alcohol,  # test model
  good ~ . - quality - drinkable,  #  full model
  data = wine,
  family = binomial(link="logit")
)
```
Set seed for reproducible results
```{r}
set.seed(1234)
```
## Number of simulations
```{r}
S <- 10^4
```
Set `X` and `y` variables
```{r}
# X=cbind(rep(1,nrow(wine)), wine$volatile_acidity, wine$total_sulfur_dioxide,
#          wine$sulphates, wine$alcohol)

X = cbind(rep(1, nrow(wine)), wine$fixed_acidity, wine$volatile_acidity, wine$citric_acid,
          wine$residual_sugar, wine$chlorides, wine$free_sulfur_dioxide, wine$total_sulfur_dioxide,
          wine$density, wine$pH, wine$sulphates, wine$alcohol)

y <- wine$good[1]
```
## Initialisations for the coefficients

The first initialisation contains the MLE estimates for each coefficient. The remaining three initialisations are randomly selected values between 500 and 1000 from the uniform distribution $runif(48, 500, 1000)$.
```{r}
# init <- matrix(data=c(runif(20, min = 100, max = 200)), nrow=4, ncol=5, byrow = T)  # for testing
init <- matrix(data=c(runif(48, min = 100, max = 200)), nrow=4, ncol=12, byrow = T)
```
## Run a Metropolis-Hastings algorithm
```{r}
# First initialisation
beta_mat_init1 <- matrix(NA, nrow = S, ncol = ncol(X))
k <- ncol(beta_mat_init1)
#beta_mat_init1[1,] <- init[1,]
beta_mat_init1[1,] <- as.numeric(coefficients(fit))  # initialise with MLE of each coefficient
acc <- 0
for (iter in 2:S) {
  # simulate all (k) values using previous value of beta as mean and set sd
  beta_star <- rnorm(k, beta_mat_init1[iter-1,], 5)
  # compute target distribution for proposed value
  newpost = lpost.LR(beta_star, X, y)
  # compute target distribution for old value
  oldpost = lpost.LR(beta_mat_init1[iter-1,], X, y)  # symmetric dist => no ratio computed
  
  # acceptance step, in log scale
  if (runif(1,0,1) > exp(newpost - oldpost)) {
    # chain doesn't move
    beta_mat_init1[iter,] = beta_mat_init1[iter-1,]
  } else {
    # add to chain and add 1 to counter
    beta_mat_init1[iter,] = beta_star
    acc=acc + 1
  }
  #if (iter%%1000 == 0) {print(c(iter, acc/iter))}
}
print(c(iter, acc/iter))

# Second initialisation
beta_mat_init2 <- matrix(NA, nrow = S, ncol = ncol(X))
k <- ncol(beta_mat_init2)
beta_mat_init2[1,] <- init[2,]
acc <- 0
for(iter in 2:S){
  beta_star <- rnorm(k, beta_mat_init2[iter-1,], 5)
  newpost = lpost.LR(beta_star, X, y)
  oldpost = lpost.LR(beta_mat_init2[iter-1,], X, y)
  if (runif(1,0,1) > exp(newpost - oldpost)) {
    beta_mat_init2[iter,] = beta_mat_init2[iter-1,]
  } else {
    beta_mat_init2[iter,] = beta_star
    acc=acc + 1
  }
  #if (iter%%1000==0) {print(c(iter,acc/iter))}
}
print(c(iter, acc/iter))

# Third initialisation
beta_mat_init3 <- matrix(NA,nrow=S,ncol=ncol(X))
k <- ncol(beta_mat_init3)
beta_mat_init3[1,] <- init[3,]
acc <- 0
for(iter in 2:S){
  beta_star <- rnorm(k,beta_mat_init3[iter-1,], 5)
  newpost=lpost.LR(beta_star,X,y)
  oldpost=lpost.LR(beta_mat_init3[iter-1,],X,y)
  if(runif(1,0,1)>exp(newpost-oldpost)){
    beta_mat_init3[iter,]=beta_mat_init3[iter-1,]
  } else{
    beta_mat_init3[iter,]=beta_star
    acc=acc+1
  }
  #if(iter%%1000==0){print(c(iter,acc/iter))}
}
print(c(iter, acc/iter))

# Forth initialisation
beta_mat_init4 <- matrix(NA,nrow=S,ncol=ncol(X))
k <- ncol(beta_mat_init4)
beta_mat_init4[1,] <- init[4,]
acc <- 0
for(iter in 2:S){
  beta_star <- rnorm(k,beta_mat_init4[iter-1,], 5)
  newpost=lpost.LR(beta_star,X,y)
  oldpost=lpost.LR(beta_mat_init4[iter-1,],X,y)
  if(runif(1,0,1)>exp(newpost-oldpost)){
    beta_mat_init4[iter,]=beta_mat_init4[iter-1,]
  } else{
    beta_mat_init4[iter,]=beta_star
    acc=acc+1
  }
  #if(iter%%1000==0){print(c(iter,acc/iter))}
}
print(c(iter, acc/iter))
```
Plot the chains for each coefficient (same plot)

```{r}
par(mfrow=c(3,2))
plot(beta_mat_init1[,1], type="l", col="magenta", ylab=expression(beta[0]))
lines(beta_mat_init2[,1], type="l", col="green")
lines(beta_mat_init3[,1], type="l", col="red")
lines(beta_mat_init4[,1], type="l", col="blue")
abline(h=fit$coefficients[1],col="red",lty=2)

plot(beta_mat_init1[,2], type="l", col="magenta", ylab=expression(beta[1]))
lines(beta_mat_init2[,2], type="l", col="green")
lines(beta_mat_init3[,2], type="l", col="red")
lines(beta_mat_init4[,2], type="l", col="blue")
abline(h=fit$coefficients[2],col="red",lty=2)

plot(beta_mat_init1[,3], type="l", col="magenta", ylab=expression(beta[2]))
lines(beta_mat_init2[,3], type="l", col="green")
lines(beta_mat_init3[,3], type="l", col="red")
lines(beta_mat_init4[,3], type="l", col="blue")
abline(h=fit$coefficients[3],col="red",lty=2)

plot(beta_mat_init1[,4], type="l", col="magenta", ylab=expression(beta[3]))
lines(beta_mat_init2[,4], type="l", col="green")
lines(beta_mat_init3[,4], type="l", col="red")
lines(beta_mat_init4[,4], type="l", col="blue")
abline(h=fit$coefficients[4],col="red",lty=2)

plot(beta_mat_init1[,5], type="l", col="magenta", ylab=expression(beta[4]))
lines(beta_mat_init2[,5], type="l", col="green")
lines(beta_mat_init3[,5], type="l", col="red")
lines(beta_mat_init4[,5], type="l", col="blue")
abline(h=fit$coefficients[5],col="red",lty=2)

plot(beta_mat_init1[,6], type="l", col="magenta", ylab=expression(beta[5]))
lines(beta_mat_init2[,6], type="l", col="green")
lines(beta_mat_init3[,6], type="l", col="red")
lines(beta_mat_init4[,6], type="l", col="blue")
abline(h=fit$coefficients[6],col="red",lty=2)
```
## Comment

* varied standard deviation of the prior in the posterior distribution function (0.5, 1, 5, 10, 20, 50)
* varied standard deviation of the Metropolis-Hastings algorithm (0.5, 1, 5, 10, 20, 50)
* varied the random initialised values (+/- 100, 0-1, 0-10, 0-100, 10-100, 100-200, 500-1000)
```{r}
fit$coefficients[1:6]
```

$\beta_0$ (`Intercept`: 242.7625): No convergence with MLE.  All four chains do converge to a value around zero.

$\beta_1$ (`fixed_acidity`: 0.275): First chain containing MLE converges to a value close but slightly under the coefficient.  The other three chains containing random initialisations converge to a value approximately 20-30.

$\beta_2$ (`volative_acidity`: -2.581): All four chains converge to a value close to the coefficient.

$\beta_3$ (`citric_acid`: 0.568): All four chains converge to a value close to the coefficient.

$\beta_4$ (`residual_sugar`: 0.239): All four chains converge to a value close to the coefficient.

$\beta_5$ (`chlorides`: -8.816): All four chains converge to a value close to the coefficient.

```{r}
par(mfrow=c(3,2))

plot(beta_mat_init1[,7], type="l", col="magenta", ylab=expression(beta[6]))
lines(beta_mat_init2[,7], type="l", col="green")
lines(beta_mat_init3[,7], type="l", col="red")
lines(beta_mat_init4[,7], type="l", col="blue")
abline(h=fit$coefficients[7],col="red",lty=2)

plot(beta_mat_init1[,8], type="l", col="magenta", ylab=expression(beta[7]))
lines(beta_mat_init2[,8], type="l", col="green")
lines(beta_mat_init3[,8], type="l", col="red")
lines(beta_mat_init4[,8], type="l", col="blue")
abline(h=fit$coefficients[8],col="red",lty=2)

plot(beta_mat_init1[,9], type="l", col="magenta", ylab=expression(beta[8]))
lines(beta_mat_init2[,9], type="l", col="green")
lines(beta_mat_init3[,9], type="l", col="red")
lines(beta_mat_init4[,9], type="l", col="blue")
abline(h=fit$coefficients[9],col="red",lty=2)

plot(beta_mat_init1[,10], type="l", col="magenta", ylab=expression(beta[9]))
lines(beta_mat_init2[,10], type="l", col="green")
lines(beta_mat_init3[,10], type="l", col="red")
lines(beta_mat_init4[,10], type="l", col="blue")
abline(h=fit$coefficients[10],col="red",lty=2)

plot(beta_mat_init1[,11], type="l", col="magenta", ylab=expression(beta[10]))
lines(beta_mat_init2[,11], type="l", col="green")
lines(beta_mat_init3[,11], type="l", col="red")
lines(beta_mat_init4[,11], type="l", col="blue")
abline(h=fit$coefficients[11],col="red",lty=2)

plot(beta_mat_init1[,12], type="l", col="magenta", ylab=expression(beta[11]))
lines(beta_mat_init2[,12], type="l", col="green")
lines(beta_mat_init3[,12], type="l", col="red")
lines(beta_mat_init4[,12], type="l", col="blue")
abline(h=fit$coefficients[12],col="red",lty=2)
```
## Comment
```{r}
fit$coefficients[7:12]
```

$\beta_6$ (`free_sulfur_dioxide`: 0.011): First chain containing MLE converges to a value close but slightly under the coefficient.  The other three chains containing random initialisations converge to a value approximately 20-30.

$\beta_7$ (`total_sulfur_dioxide`: -0.017): First chain converges to a value just below coefficient but chains 2-4 don't converge to coefficient.

$\beta_8$ (`density`: -257.798): First chain converges to a value at about the coefficient but chains 2-4 converge to a value slightly above the coefficient

$\beta_9$ (`pH`: 0.224): All four chains converge to a value close to the coefficient.

$\beta_{10}$ (`sulphates`: 3.750): All four chains converge to a value close to the coefficient.

$\beta_{11}$ (`alcohol`: 0.753): First chain converges to a value at about the coefficient but chains 2-4 converge to a value above the coefficient


# 7. Posterior Predictive Distribution
Approximate the posterior predictive distribution of an unobserved variable characterised by the following values for each covariate:
```{r}
fixed_acidity <- 7.5
volatile_acidity <- 0.6
citric_acid <- 0.0
residual_sugar <- 1.7
chlorides <- 0.085
free_sulfur_dioxide <- 5
total_sulfur_dioxide <- 45
density <- 0.9965
pH <- 3.4
sulphates <- 0.63
alcohol <- 12
```


```{r}
S <- 20000
beta_mat2 <- matrix(NA, nrow = S, ncol = ncol(X))
beta_mat2[1,] <- as.numeric(coefficients(fit))

y_new <- c(1)
# x_new <- c(1, 5, 0.63, 12, 0.6)  # for testing
x_new <- c(1, fixed_acidity, volatile_acidity, citric_acid, residual_sugar,
           chlorides, free_sulfur_dioxide, total_sulfur_dioxide, density,
           pH, sulphates, alcohol)
```

new model
```{r}
library(mvtnorm)

# prediction

Omega_prop <- solve(t(X) %*% X)
k <- ncol(beta_mat2)
acc <- 0
for(iter in 2:S)
{
  # 1. Propose a new set of values
  beta_star <- rmvnorm(1, beta_mat2[iter-1,], 1.5 * Omega_prop)
  
  # 2. Compute the posterior density on the proposed value and on the old value  
  newpost=lpost.LR(t(beta_star), X, y)
  oldpost=lpost.LR(matrix(beta_mat2[iter-1,], ncol=1), X, y)
  
  # 3. Acceptance step
  if (runif(1, 0, 1) > exp(newpost - oldpost)) {
    beta_mat2[iter,] = beta_mat2[iter-1,]
  } else {
    beta_mat2[iter,] = beta_star
    acc = acc + 1
  }
  # 4. Print the stage of the chain
  if (iter%%1000 == 0){ print(c(iter, acc/iter)) }
  
  # 5. Prediction 
  p_new <- exp(sum(beta_mat2[iter,] * x_new) ) / (1 + exp(sum(beta_mat2[iter,] * x_new) ))
  y_new[iter] <- rbinom(1,1,prob=p_new)
}
```

Plots

```{r}
par(mfrow=c(3,2))
plot(beta_mat2[,1],type="l", ylab=expression(beta[0]))
abline(h=fit$coefficients[1],col="red",lty=2)
plot(beta_mat2[,2],type="l", ylab=expression(beta[1]))
abline(h=fit$coefficients[2],col="red",lty=2)
plot(beta_mat2[,3],type="l", ylab=expression(beta[2]))
abline(h=fit$coefficients[3],col="red",lty=2)
plot(beta_mat2[,4],type="l", ylab=expression(beta[3]))
abline(h=fit$coefficients[4],col="red",lty=2)
plot(beta_mat2[,5],type="l", ylab=expression(beta[4]))
abline(h=fit$coefficients[5],col="red",lty=2)
plot(beta_mat2[,6],type="l", ylab=expression(beta[5]))
abline(h=fit$coefficients[6],col="red",lty=2)
```

Plots


```{r}
par(mfrow=c(3,2))
plot(beta_mat2[,7],type="l", ylab=expression(beta[6]))
abline(h=fit$coefficients[7],col="red",lty=2)
plot(beta_mat2[,8],type="l", ylab=expression(beta[7]))
abline(h=fit$coefficients[8],col="red",lty=2)
plot(beta_mat2[,9],type="l", ylab=expression(beta[8]))
abline(h=fit$coefficients[9],col="red",lty=2)
plot(beta_mat2[,10],type="l", ylab=expression(beta[9]))
abline(h=fit$coefficients[10],col="red",lty=2)
plot(beta_mat2[,11],type="l", ylab=expression(beta[10]))
abline(h=fit$coefficients[11],col="red",lty=2)
plot(beta_mat2[,12],type="l", ylab=expression(beta[11]))
abline(h=fit$coefficients[12],col="red",lty=2)
```


```{r}
table(y_new[15000:20000])
```
# 8. metrop() analysis of Q6
Implementation of the `metrop()` function was through the [MCMC Package Example, Charles J. Geyer](https://cran.r-project.org/web/packages/mcmc/vignettes/demo.pdf)


```{r}
library(mcmc)
# out <- glm(wine$quality ~ wine$volatile_acidity + wine$total_sulfur_dioxide + wine$sulphates + wine$alcohol, data=wine, family=binomial, x=TRUE)
# summary(out)

# full model
out <- glm(good ~ . - quality - drinkable, data=wine, family=binomial, x=TRUE)
```

```{r}
lupost_factory <- function(x, y) function(beta) {
  eta <- as.numeric(x %*% beta)
  logp <- ifelse(eta < 0, eta - log1p(exp(eta)), - log1p(exp(- eta)))
  logq <- ifelse(eta < 0, - log1p(exp(eta)), - eta - log1p(exp(- eta)))
  logl <- sum(logp[y == 1]) + sum(logq[y == 0])
  return(logl - sum(beta^2) / 8)
}
lupost <- lupost_factory(out$x, out$y)
```


```{r}
set.seed(1234)
beta.init <- as.numeric(coefficients(out))
out <- metrop(lupost, beta.init, 1e3)

names(out)
```
Look at the acceptance rate.

```{r}
out$accept
```
This is very low, so we can adjust the scale parameter to find an acceptance rate around 20%.
```{r}
out <- metrop(out, scale = 0.0075)
out$accept
```


```{r}
par(mfrow=c(3,2))
plot(ts(out$batch)[,1], ylab=expression(beta[0]))
abline(h=fit$coefficients[1],col="red",lty=2)
plot(ts(out$batch)[,2], ylab=expression(beta[1]))
abline(h=fit$coefficients[2],col="red",lty=2)
plot(ts(out$batch)[,3], ylab=expression(beta[2]))
abline(h=fit$coefficients[3],col="red",lty=2)
plot(ts(out$batch)[,4], ylab=expression(beta[3]))
abline(h=fit$coefficients[4],col="red",lty=2)
plot(ts(out$batch)[,5], ylab=expression(beta[4]))
abline(h=fit$coefficients[5],col="red",lty=2)
plot(ts(out$batch)[,6], ylab=expression(beta[5]))
abline(h=fit$coefficients[6],col="red",lty=2)
```



```{r}
par(mfrow=c(3,2))
plot(ts(out$batch)[,7], ylab=expression(beta[0]))
abline(h=fit$coefficients[7],col="red",lty=2)
plot(ts(out$batch)[,8], ylab=expression(beta[1]))
abline(h=fit$coefficients[8],col="red",lty=2)
plot(ts(out$batch)[,9], ylab=expression(beta[2]))
abline(h=fit$coefficients[9],col="red",lty=2)
plot(ts(out$batch)[,10], ylab=expression(beta[3]))
abline(h=fit$coefficients[10],col="red",lty=2)
plot(ts(out$batch)[,11], ylab=expression(beta[4]))
abline(h=fit$coefficients[11],col="red",lty=2)
plot(ts(out$batch)[,12], ylab=expression(beta[5]))
abline(h=fit$coefficients[12],col="red",lty=2)
```
## VISUAL COMPARISON

